---
title: "Duc Ngo Test"
author: "Duc Ngo"
output: html_document
---
hey
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)         # for reading in data, graphing, and cleaning
library(tidymodels)        # for modeling ... tidily
library(usemodels)         # for suggesting step_XXX() functions
library(glmnet)            # for regularized regression, including LASSO
library(naniar)            # for examining missing values (NAs)
library(lubridate)         # for date manipulation
library(moderndive)        # for King County housing data
library(vip)               # for variable importance plots
library(rmarkdown)         # for paged tables
library(dplyr)
library(janitor)
# Load packages
library(bayesrules)
library(tidyverse)
library(rstanarm)
library(broom.mixed)
library(tidybayes)
library(bayesplot)
library(ggplot2)
theme_set(theme_minimal()) # my favorite ggplot2 theme :)
```

```{r}
library(readr)
final_dataset <- read_csv("FINALDATASET.csv")
```

```{r}
top_50 <- final_dataset %>% 
  filter(YEAR == 2021) %>% 
  arrange(EARNINGS)  %>% 
  tail(50)

top_50 <- (top_50$COMPANY)

top_50 <- final_dataset %>% 
    filter(COMPANY %in% top_50)

write.csv(top_50, "TOP_50_COMPANIES_update.csv")
```

```{r}
top_50_update <- top_50 %>% 
  drop_na()
```

### First simple model: without the hiarchical model- predicting the earnings using only the last year's earnings and sales.

```{r}
# Simulate the posterior
easy_model <- stan_glm(
  Earnings_next_year ~ EARNINGS + SALES, data = top_50_update,
  family = gaussian,
  prior_PD = FALSE,
  chains = 4, iter = 5000*2, seed = 84735, refresh = 0)
```


```{r}
# Plot 200 posterior plausible model lines
top_50_update %>%
  add_fitted_draws(easy_model, n = 200) %>%
  ggplot(aes(x = Earnings_next_year, y = EARNINGS)) +
    geom_line(aes(y = .value, group = .draw), alpha = 0.15) + 
    geom_point(data = top_50_update)
```


```{r}
tidy(easy_model, conf.int = TRUE, conf.level = 0.95)
```

```{r}
tidy(easy_model, effects = "fixed", conf.int = TRUE, conf.level = 0.80)
```

```{r}
pp_check(easy_model)
```



### Hiarchical model: 


#### Part 1: Clean the NA value first

```{r}
final_dataset2 <- final_dataset %>% 
  drop_na()
```




```{r}
# Simulate the prior model
model_1_prior_with_hiar <- stan_glmer(
  Earnings_next_year ~ EARNINGS + (1 | COMPANY), data = top_50_update, 
  family = gaussian,
  chains = 4, iter = 5000*2, seed = 84735, 
  prior_PD = TRUE)
```


tidy(model_1_prior_with_hiar, conf.int = TRUE)

model_1_post_with_hiar <- stan_glmer(
  EARNINGS ~ SALES + (1 | COMPANY), data = top50Companies, 
  family = gaussian,
  chains = 4, iter = 5000*2, seed = 84735, 
  prior_PD = FALSE)

tidy(model_1_post_with_hiar, conf.int = TRUE)

```{r}
# Check out the prior specifications
prior_summary(model_1_prior_with_hiar)
```


```{r}
final_dataset2%>%
  add_predicted_draws(model_1_prior_with_hiar, n = 100) %>%
  ggplot(aes(x = Earnings_next_year)) +
    geom_density(aes(x = .prediction, group = .draw)) 
```

```{r}
model_1 <- update(model_1_prior_with_hiar, prior_PD = FALSE, refresh = 0)
pp_check(model_1)
```

Check the accuracy of the model: 

```{r}
tidy(model_1, effects = "fixed", conf.int = TRUE, conf.level = 0.80)
```

```{r}
tidy(model_1, effects = "ran_pars")
```


#### Maybe better model?

```{r}
# Simulate the prior model
model_2_more_complex_with_hiar <- stan_glmer(
  Earnings_next_year ~ EARNINGS + SALES +  (1 | COMPANY), data = top_50_update, 
  family = gaussian,
  chains = 4, iter = 5000*2, seed = 84735, 
  prior_PD = FALSE)
```

```{r}
pp_check(model_2_more_complex_with_hiar)
```














