---
title: "Untitled"
author: "NicholasDi"
date: "11/26/2021"
output: html_document
---

```{r setup, include=FALSE}
#Setup
library(rvest)
library(dplyr)
library(lubridate)
library(stringr)
library(tidymodels)
library(bayesmodels)
library(modeltime)
library(tidyverse)
library(timetk)
interactive = FALSE
```

```{r}
m750 <- m4_monthly %>% 
  filter(id == "M750")
m750 %>%
  plot_time_series(date, value, .interactive = interactive)
```

Bayesmodels and Modeltime models (e.g. sarima_reg() and arima_reg()) are created with a date or date time feature in the model. You will see that most models include a formula like fit(value ~ date, data).

Parsnip models (e.g. linear_reg()) typically should not have date features, but may contain derivatives of dates (e.g. month, year, etc). You will often see formulas like fit(value ~ as.numeric(date) + month(date), data).

```{r}
#Split Data 80/20
splits <- initial_time_split(m750, prop = 0.9)
```

```{r}
#Model 1: Arima ----
model_fit_arima<- arima_reg(non_seasonal_ar = 0,
                            non_seasonal_differences = 1,
                            non_seasonal_ma = 1,
                            seasonal_period = 12,
                            seasonal_ar = 0,
                            seasonal_differences = 1,
                            seasonal_ma = 1) %>%
    set_engine(engine = "arima") %>%
    fit(value ~ date, data = training(splits))
```

```{r}
model_fit_arima_bayes<- sarima_reg(non_seasonal_ar = 0,
                                  non_seasonal_differences = 1,
                                  non_seasonal_ma = 1,
                                  seasonal_period = 12,
                                  seasonal_ar = 0,
                                  seasonal_differences = 1,
                                  seasonal_ma = 1,
                                  pred_seed = 100) %>%
    set_engine(engine = "stan") %>%
    fit(value ~ date, data = training(splits))
```

```{r}
plot(model_fit_arima_bayes$fit$models$model_1)
```

```{r}
data <- read.csv("TOP_50_COMPANIES_update.csv")
APPL <- data %>% 
  filter(COMPANY == 'AMZN') %>% 
  select(EARNINGS) %>% 
  arrange(EARNINGS)

vector <- APPL$EARNINGS
```
```{r}
model_fit_naive <- random_walk_reg(seasonal_random_walk = TRUE, seasonal_period = 12) %>%
                   set_engine("stan") %>%
                   fit(value ~ date + month(date), data = training(splits))
```

```{r}
plot(model_fit_naive$fit$models$model_1)
```

```{r}
models_tbl <- modeltime_table(
    model_fit_arima,
    model_fit_arima_bayes,
    model_fit_naive
)
```





```{r}
# link = "https://www.snowpak.com/usa/best-ski-resorts"
# page = read_html(link)
# name = page %>% html_nodes('h2') %>% html_text()
# dates = page %>% html_nodes('.jHpJNR li') %>% html_text()
# 
# #Clean Dates
# dates <- gsub("\\*.*","", dates)
# name <- gsub(".*\\.","", name)
# 
# #Data Cleaning - Resort Name
# name = as.data.frame(name)
# name$index <- 1:nrow(name)
# 
# #Data Cleaning Open and Close Dates
# dates = as.data.frame(dates)
# dates$index <- rep(1:(nrow(dates)/2), each = 2)
# dates <- dates %>% 
#   group_by(index) %>% 
#   summarise(team_no_nerged = paste(dates, collapse = "-"))
# 
# #Joining
# data <- left_join(name, dates)
# data <- data %>% 
#   select(-'index')
```

