---
title: "Nick-Models"
author: "NicholasDi"
date: "12/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Setting Up
```{r}
library(tidyverse)         # for reading in data, graphing, and cleaning
library(tidymodels)        # for modeling ... tidily
library(usemodels)         # for suggesting step_XXX() functions
library(glmnet)            # for regularized regression, including LASSO
library(naniar)            # for examining missing values (NAs)
library(lubridate)         # for date manipulation
library(moderndive)        # for King County housing data
library(vip)               # for variable importance plots
library(rmarkdown)         # for paged tables
library(dplyr)
library(janitor)
library(ggplot2)
library(dplyr)
library(bayesforecast)    # Bayes Forecasting (Playing around) 
library(bayesrules)
library(tidyverse)
library(rstanarm)
library(broom.mixed)
library(tidybayes)
library(bayesplot)
library(ggplot2)
library(tidymodels)
library(bayesmodels)
library(modeltime)
library(tidyverse)
library(timetk)
library(lubridate)
# This toggles plots from plotly (interactive) to ggplot (static)
interactive <- FALSE
```

### Dataset: 

```{r, echo=FALSE}
#Loading Dataset 
randomCompany <- read_csv("RandomCompnay.csv")
```

```{r}
#Scaling Variables of Interest
data <- randomCompany %>% 
  mutate(EARNINGS_Scaled = EARNINGS/1000000000,
         CASH_Scaled = CASH/1000000000,
         MARKET.CAP_Scaled = MARKET.CAP/1000000000,
         Earnings_next_year_Scaled = Earnings_next_year/1000000000,
         SALES_Scaled = SALES/1000000000)

#Adding Lagged Variables
data <- data %>% 
  group_by(COMPANY) %>%
  mutate(EARNINGS_1_YEAR_AGO = lead(EARNINGS_Scaled, n = 1), 
         EARNINGS_2_YEAR_AGO = lead(EARNINGS_Scaled, n = 2),
         EARNINGS_3_YEAR_AGO = lead(EARNINGS_Scaled, n = 3),
         EARNINGS_4_YEAR_AGO = lead(EARNINGS_Scaled, n = 4)) %>% 
  mutate(SALES_1_YEAR_AGO = lead(SALES_Scaled, n = 1),
         SALES_2_YEAR_AGO = lead(SALES_Scaled, n = 2),
         SALES_3_YEAR_AGO = lead(SALES_Scaled, n = 3),
         SALES_4_YEAR_AGO = lead(SALES_Scaled, n = 4))
data <- data[,-1] 

         
```

```{r}
head(data)
```
# Model 1 

This model is a regular simple normal regression, as we are not taking advantage of thei grouped strucuture of our data. We are having a complete pooled regression here. 

```{r}
model_1 <- stan_glm(
  Earnings_next_year_Scaled ~ EARNINGS_Scaled, data = data,
  family = gaussian,
  prior_PD = FALSE,
  chains = 4, iter = 5000*2, seed = 84735, refresh = 0)
```

```{r}
#Check how our priors were tuned
prior_summary(model_1)
mean(data$EARNINGS_Scaled)
```

## Model Notation 

Where Y is Earnings next year in billions and X is Earnings in the current year in billions. 

Below is our model notation with "specified" priors 
$$\begin{split}
Y_i | \beta_0, \beta_1, \sigma & \stackrel{ind}{\sim} N(\mu_i, \sigma^2) \;\; \text{ where } \mu_i = \beta_0 + \beta_1 X_i\\
\beta_{0c} & \sim N(1.1, 2.5^2) \\
\beta_1    & \sim N(0, 2.5^2) \\
\sigma     & \sim \text{Exp}(1) \\
\end{split}$$

Below is our model notation with adjusted priors

$$\begin{split}
Y_i | \beta_0, \beta_1, \sigma & \stackrel{ind}{\sim} N(\mu_i, \sigma^2) \;\; \text{ where } \mu_i = \beta_0 + \beta_1 X_i\\
\beta_{0c} & \sim N(1.1, 6.8^2) \\
\beta_1    & \sim N(0, 2.5^2) \\
\sigma     & \sim \text{Exp}(0.37) \\
\end{split}$$

```{r}
#Analyzing Coefficients 
tidy(model_1, effects = "fixed", conf.int = TRUE, conf.level = 0.80)
```
```{r}
pp_check(model_1)
```

## Model 1 Using Previous Years

```{r}
model_1_2 <- stan_glm(
  Earnings_next_year_Scaled ~ EARNINGS_Scaled + EARNINGS_1_YEAR_AGO + EARNINGS_2_YEAR_AGO, data = data,
  family = gaussian,
  prior_PD = FALSE,
  chains = 4, iter = 5000*2, seed = 84735, refresh = 0)
```

```{r}
prior_summary(model_1_2)
```


```{r}
pp_check(model_1_2)
```



#Model 2

Similar to the above regression, below we regress earnings next year with a sector dummy variable. 

```{r}
model_2 <- stan_glm(
  Earnings_next_year_Scaled ~ Sector, data = data,
  family = gaussian,
  prior_PD = FALSE,
  chains = 4, iter = 5000*2, seed = 84735, refresh = 0)
```

```{r}
#checking to see how priors are tuned
prior_summary(model_2)
```

## Model Notation 

Below we regress Earnings Next Year depending on Sector. We have a total of 11 sectors, thus there will be 10 coefficents besides the intercept. Below I plug in our specified prior values.

$$\begin{split}
Y_i | \beta_0, \beta_1, \sigma & \stackrel{ind}{\sim} N(\mu_i, \sigma^2) \;\; \text{ where } \mu_i = \beta_0 + \beta_1 (Sector2)X_i + \beta_2 (Sector3)X_i...\\
\beta_{0c} & \sim N(1.1, 2.5^2) \\
\beta_1    & \sim N(0, 2.5^2) \\
\beta_2    & \sim N(0, 2.5^2) \\
\beta_3    & \sim N(0, 2.5^2) \\
.\\
.\\
.\\
\sigma     & \sim \text{Exp}(1) \\
\end{split}$$

Here are our Adjusted prior values. 

$$\begin{split}
Y_i | \beta_0, \beta_1, \sigma & \stackrel{ind}{\sim} N(\mu_i, \sigma^2) \;\; \text{ where } \mu_i = \beta_0 + \beta_1 (Sector2)X_i + \beta_2 (Sector3)X_i...\\
\beta_{0c} & \sim N(1.1, 6.8^2) \\
\beta_1    & \sim N(0, 18.77^2) \\
\beta_2    & \sim N(0, 63.24^2) \\
\beta_3    & \sim N(0, 30.59^2) \\
.\\
.\\
.\\
\sigma     & \sim \text{Exp}(0.37) \\
\end{split}$$

```{r}
tidy(model_2, effects = "fixed", conf.int = TRUE, conf.level = 0.80)
```
```{r}
pp_check(model_2)
```

The reference coefficient is "Communication Services" 

#Model 3 - Using Heir Structure

Now we will use a regression where we will utilize the structure of our data. 

```{r}
model_3_prior_with_hiar <- stan_glmer(
  Earnings_next_year_Scaled ~ EARNINGS_Scaled + (1 | COMPANY), data = data, 
  family = gaussian,
  chains = 4, iter = 5000*2, seed = 84735, 
  prior_PD = TRUE)
```

```{r}
prior_summary(model_3_prior_with_hiar)
```

##Model Notation 

$$\begin{split}
\text{Relationship within company:} & \\
Y_{ij} | \beta_{0j}, \beta_1, \sigma_y 
& \sim N(\mu_{ij}, \sigma_y^2) \;\; \text{ where } \mu_{ij} = \beta_{0j} + \beta_1 X_{ij} \\
& \\
\text{Variability between companies:} & \\
\beta_{0j} & \stackrel{ind}{\sim} N(\beta_0, \sigma_0^2) \\
& \\
\text{Prior information on Globals} & \\
\beta_{0c} & \sim N(m_0, s_0^2) \\
\beta_1 & \sim N(m_1, s_1^2) \\
\sigma_y  & \sim \text{Exp}(l_y) \\
\sigma_0  & \sim \text{Exp}(l_0) \\
\end{split}$$

```{r}
tidy(model_3_prior_with_hiar, effects = "fixed", conf.int = TRUE, conf.level = 0.80)
```

```{r}
pp_check(model_3_prior_with_hiar)
```

## Heirc Models Using Previous Years Data

```{r}
model_3_prior_with_hiar_2 <- stan_glmer(
  Earnings_next_year_Scaled ~ EARNINGS_Scaled + EARNINGS_1_YEAR_AGO + EARNINGS_2_YEAR_AGO + (1 | COMPANY), data = data, 
  family = gaussian,
  chains = 4, iter = 5000*2, seed = 84735, 
  prior_PD = TRUE)
```

```{r}
pp_check(model_3_prior_with_hiar_2)
```


### Bayesian Forecast: 

```{r}
library(bayesforecast)
```

```{r}
ts_model <- function(company) {
  newdata <- data %>% 
    filter(COMPANY == company) %>% 
    select(EARNINGS_Scaled) %>% 
    arrange(EARNINGS_Scaled)
  
  vector <- company$EARNINGS_Scaled
  
  myts <- ts(vector, start=c(1999), end=c(2021), frequency=1)

  sf1 = stan_sarima(ts = myts,order = c(1,1,1),seasonal = c(1,1,1),
                  prior_mu0 = student(mu = 0,sd = 1,df = 7))
  
}

ts_model(company = "AMZN")
```

```{r}
AMZN <- data %>% 
  filter(COMPANY == 'AMZN') %>% 
  select(EARNINGS_Scaled) %>% 
  arrange(EARNINGS_Scaled)

vector <- AMZN$EARNINGS_Scaled

myts <- ts(vector, start=c(1999), end=c(2021), frequency=1)

sf1 = stan_sarima(ts = myts,order = c(1,1,1),seasonal = c(1,1,1),
                  prior_mu0 = student(mu = 0,sd = 1,df = 7))
```

```{r}
data %>% 
  select(EARNINGS)
```

```{r}
AMZN <- data %>% 
  filter(COMPANY == 'AMZN') %>% 
  select(EARNINGS) 
  #arrange(EARNINGS_Scaled)
```

```{r}
sf1
```

```{r fig3}
check_residuals(sf1)
autoplot(forecast(object = sf1,h = 12))
```


